Does It Scale? Static Output Approximation of PHP Web Applications Using Symbolic Execution

### 2017-01-03 paper.tex
3.1.2016 es schneit

@@

@@

\lstset{
language=PHP,
basicstyle=\ttfamily\footnotesize,
basewidth={0.5em,0.4em}, % for other columns settings
%columns=fullflexible,
numbers=left,
numberstyle=\tiny,
frame=single,
showstringspaces=false,
xleftmargin=2em,
framexleftmargin=1.5em
}

@@

\begin{document}
\lstset{
	language=PHP,
	basicstyle=\footnotesize\ttfamily,
    commentstyle = \color{gray},
    extendedchars = \true,
    inputencoding = utf8x,
    keepspaces = true,
    keywordstyle = \bfseries
}

@@

\emph{Symbolic execution} is a static analysis technique between program testing
and program proving. For a program, symbolic, i.e., abstract, yet fix inputs are
applied instead of concrete inputs \cite{Darringer1978,King1976}. The underlying
concept is to map program input (symbolically) to program output. Different classes of inputs may
result in variational output, representing dependencies of output and input.
As symbolic execution of a program returns an input-output-mapping for symbolic inputs, this analysis technique helps unfold the \emph{staged} nature of web applications: Since every feasible path is executed all possible output is contained in the symbolic output, and all dynamically generated variational parts of the web application can be analyzed.

@@

\section{State Of the Art}
In this section we present the motivation for building tools that address the \textcolor{blue}{development, quality and xxx} of dynamic web applications, but require an accurate static output approximation. All tools proposed so far \cite{Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140,Nguyen:2011:AFH:2190078.2190142} leverage a representation of all possible HTML client page output. Since any variant can be analyzed in particular, this enables tool support addressing the web application as a whole despite its \emph{staged} nature, where parts of the web application (client-side output) is just generated dynamically.

The work-flow for all these tools is illustrated in Figure \ref{fig:approach}, where prior to any analysis, the web application is symbolically executed and all client page output variants are parsed. To illustrate the concept of symbolic execution for this purpose consider the following example:

\begin{figure}[t!]
	\centering
	\includegraphics[width=0.45\textwidth]{images-paper/approach}

@@

\begin{lstlisting}
<?php

echo 'Hello World!';
\end{lstlisting}

Based on this output representation, the so-called D-Model, subsequent analyses can be conducted:

@@

\item \emph{WebSlice}: Bener bener
\item \emph{Varis}: Bener IDE bener
\end{compactitem}

\textcolor{blue}{\emph{D. Tool evaluation critique}
\begin{compactitem}
	\item Limitations/Assumptions
	\item Too small sample
\end{compactitem}
}

@@

\subsection{Measuring Approximation Success} \label{heuristic}
For the evaluation of our output approximation we require a ground truth output model to compare against.
In the absence of ground truth for the approximated output we are bound to choose a \emph{heuristic approach} in order to measure the accuracy of our aproximation. We define a string literal to be an \emph{output candidate} if there exists an execution path reaching the string literal and evantually passing it to an 
output-generating statement (e.g., \texttt{echo} or \texttt{print}).

@@

\end{compactitem}
As for for the last two cases, we limited our selection since the parser we
used did not support all language features used.

\begin{table*}[t]
\centering 
	\begin{tabular}{lrp{4cm}rlrr}
	\toprule
	\textbf{System} & \textbf{Version} & \textbf{Classification} & \textbf{SLOC} & \textbf{\#files} & \textbf{\#OCs} & \textbf{\#includes}
	\\
	\midrule
	\textsc{AddressBook} & 8.2.5.2 & 239 & 51,907 & Adress Administration  & XXX & XXX\\
	\textsc{SchoolMate} & 1.5.4 & 65 & 8,118 & School Administration  & XXX & XXX\\
	\textsc{TimeClock} & 1.04 & 63 & 20,800 & XXX  & XXX & XXX\\
	\textsc{WebChess} & 1.0.0 & 28 & 5,219 & XXX  & XXX & XXX\\
	\midrule
	\textsc{Drupal} & 7.5.0 & 125 & 52,464 & Content Management System  & XXX & XXX\\
	\textsc{phpBB} & 3.1.9 & 1,398 & 327,371 & Bulletin Board  & XXX & XXX\\
	\textsc{phpMyAdmin} & 4.6.3 & 871 & 303,582 & Database Administration  & XXX & XXX\\
	\midrule
	\textsc{Anchor} & 0.12.1 & 201 & 15,054 & Content Management System  & XXX & XXX\\
	\textsc{Kirby} & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
	\textsc{Automad} & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
	\textsc{Monstra} & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
	\textsc{Nibbleblog} & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
	\bottomrule
	\end{tabular}
	\caption{Corpus of twelve PHP systems. The file count includes files with a .php,
	.inc, .bit or .module extension.}
	\label{corpus}
\end{table*}

For the experimental analysis we symbolically executed per system each script file or entry point (files with a \texttt{.php}, \texttt{.inc}, \texttt{.bit} or \texttt{.module} extension). All following measurements are cumulated per system for each entry point.

@@

\subsubsection{What literals did we miss?}
\label{WhatLiteralsDidWeMiss?}
Our initial approach to understand missed output candidates is to find out whether their surrounding program code was not accessed/accessible, and if so, why. Of special interest are those output candidates that are located in HTML files (possibly with nested PHP), or located in a function definition. Plain HTML files do not require any further execution (of course, unless PHP scripts are embedded) and represent output candidates that just need to be included properly to be part of the output. Aside, output candidates that are part of a function definition are only missed if the corresponding function is never called at runtime.

We started classifying missed output candidates by their string literal context. As the classification statistics in Figure \ref{fig:contexts} illustrate, for almost all systems (except for \textsc{AddressBook} and \textsc{TimeClock}) not-reached output candidates had a function context. Note that \textsc{SchoolMate} was excluded from the diagram since our analysis reached all output candidates for that system.

@@

Finally, we measured for include expressions the overall reach coverage and the overall resolution success rate per system. As illustrated in Figure \ref{fig:include_coverage_results2}, except for the small-scale systems and two recent small-scale systems we did not even really reach a great portion of include expressions, although for most systems the resolution success rate was over 80 percent (see Figure \ref{fig:include_coverage_results1}). Nevertheless, one failed attempt to include a script file can result in a cascade of missing more include expressions as the include target script is not executed.

@@

    \label{fig:include_coverage_results}
    \caption{Include coverage and accuracy.}
\end{figure}

@@

The callgraph snippets in Figure \ref{fig:callgraph} illustrate the idea of detecting the functions’ access points: The nodes at the top depict callback candidates, the nodes at the bottom with labels X, Y, and Z depict function definitions containing not-reached output candidates. In Figure \ref{fig:callgraph_a} function X can only be accessed through callback candidates, as for Figure \ref{fig:callgraph_b} function Y can be partially accessed by callback candidates, whereas function Z is never accessible through callback candidates.

@@

For failed function calls to functions containing not-reached output candidates we measured how many of those functions were only, or partially accessibly through indirect calling mechanisms (see Section \ref{sec:inaccessible}). As Figure \ref{fig:output_candidate_explanation} illustrates, around 80 percent (except for \textsc{TimeClock}) all missed output candidates of each system were only accessible through callback candidates, i.e., functions that are only accessible through indirect calling mechanisms.

>>>>

\setlength\parindent{0pt}

@@

\usepackage[zerostyle=d]{newtxtt}
%\renewcommand*\familydefault{\ttfamily} %% Only if the base font of the document is to be typewriter style
\usepackage[T1]{fontenc}

@@

\lstset{
language=PHP,
basicstyle=\ttfamily\tiny,
basewidth={0.5em,0.4em}, % for other columns settings
%columns=fullflexible,
numbers=left,
numberstyle=\tiny,
frame=single,
showstringspaces=false,
xleftmargin=2em,
framexleftmargin=1.5em
}

@@

\begin{document}
\lstset{
	language=PHP,
	basicstyle=\footnotesize\ttfamily,
    commentstyle = \color{gray},
    extendedchars = \true,
    inputencoding = utf8x,
    keepspaces = true,
    keywordstyle = \bfseries,
    frame = tb
}

@@

\emph{Symbolic execution} is a static analysis technique between program testing
and program proving. For a program, symbolic, i.e., abstract, yet fixed inputs are
applied instead of concrete inputs \cite{Darringer1978,King1976}. The underlying
concept is to map program input (symbolically) to program output. Different classes of inputs may
result in variational output, representing dependencies of output and input.
As symbolic execution of a program returns an input-output-mapping for symbolic inputs, this analysis technique helps unfold the \emph{staged} nature of web applications: Since every feasible path is executed all possible output is contained in the symbolic output, and all dynamically generated variational parts of the web application can be analyzed.

@@

\section{State Of the Art}
Symbolic execution is one of integral analysis techniques used by the tools whose practicality we aim to evaluate. This section recaps the idea of symbolic execution, the tools which make use of it, and both conceptual and language-related limitations  of symbolic execution applied to PHP so far.

\subsection{Symbolic Execution}
Symbolic execution is a static analysis technique that was proposed by James C. King in 1976 \cite{King1976}. Symbolic execution allows to explore all feasible paths in a program. Based on normal program execution, the execution semantics is extended to handle symbolic values, which are abstract, but fixed; an execution becomes symbolic by introducing symbolic values as program inputs instead of concrete values \cite{King1976,Darringer1978}. These symbolic values are then propagated and used throughout the execution.

Execution starts with a plain path condition, usually a tautology. Whenever the control-flow can be split into different branches, the corresponding path conditions (guards for instance) are conjoined with the previous path condition \cite{King1976}. Additionally, the conjoined path condition can be checked for satisfiability to exclude infeasible paths. Consequently, for a given program symbolic execution computes a mapping from (symbolic) input values to output values.

% Introduction example for symbolic execution:
\begin{figure}
\begin{lstlisting}
<?php
$x = $_POST['name'];
$y = $x + 1;
\end{lstlisting}

\begin{lstlisting}
<?php
function foo($arg) {
   echo $arg;	
   foo($arg);
}
$z = foo(3);
\end{lstlisting}
\begin{lstlisting}
<?php
if ($z < 0) {
   $r = 'A';
} else {
   $r = 'B';
}
echo $r;
\end{lstlisting}
\begin{lstlisting}
<?php
while ($z > 0) {
   echo 'B';
   $z = $z - 1;
}
\end{lstlisting}
\caption{Example for symbolic execution of PHP code\label{fig:symbolic}}
\end{figure}

\subsection{Workflow And Existing Tools} 
The work-flow for all existing tools is illustrated in Figure \ref{fig:approach}. Prior to any analysis, the web application is symbolically executed which for every execution path yields its corresponding client page output variant. Next, all client page output variants are parsed and merged into a compact representation of all client page output variants. This output model, also referred to as \emph{VarDOM} (Variational Document Object Model), is similar to a DOM, yet it can contain symbolic value nodes and conditional node types to represent variants as alternatives.

% Workflow Graphic
\begin{figure}[t!]
	\centering
	\includegraphics[width=0.45\textwidth]{images-paper/approach}

@@

All tools proposed so far leverage a representation of all possible HTML client page output. Since any variant can be analyzed in particular, this enables tool support addressing the web application as a whole in spite of its staged nature, where parts of the web application (client-side output) are generated dynamically.

Based on the approximated output representation subsequent analyses can be conducted, which have been implemented in the following tools:

@@

\item \emph{WebSlice}: Program slices enable to extract and understand the impact of changes in an application. 
To consider all of a dynamic web application for program slices, client page output needs to be taken into account. WebSlice combines PHP data-flow information with an output approximation to enable program slices across different languages \cite{Nguyen:2015:CPS:2786805.2786872}.

\item \emph{Varis}: Editor services such as “jump to declaration” are nontrivial for dynamic web applications due to their staged nature. Varis provides editor services across stages for client-side code: Starting with an ouput approximation a callgraph with conditional edges is constructed, which allows navigation in client-site code (HTML, JS and CSS) although it is embedded in server-side code \cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
\end{compactitem}

\subsection{Assumptions and Limitations}
\textcolor{blue}{
\begin{compactitem}
	\item Design Decisions: Single loop and recursion execution due to conceptual limitations
	\item Conceptual Limitations: Symbolic input as form input, interaction
	\item Language-related or implementation-specific: No OOP, array support, standard lib
\end{compactitem}
}

@@

\subsection{Measuring Approximation Success} \label{heuristic}
For the evaluation of our output approximation we require a ground truth output model to compare our approximation against. Since output literals are embedded in server-side code and scattered, it is nontrivial to determine whether a string literal is part of the ouput or not. So, in the absence of ground truth for the approximated output we are bound to choose a \emph{heuristic approach} in order to measure the accuracy of our aproximation. We define a string literal to be an \emph{output candidate} if there exists an execution path reaching the string literal and evantually passing it to an 
output-generating statement (e.g., \texttt{echo} or \texttt{print}).

@@

\end{compactitem}
As for for the last two cases, we limited our selection since the parser we
used did not support all language features used. For the experimental evaluation we symbolically executed per system each script file or entry point (files with a \texttt{.php}, \texttt{.inc}, \texttt{.bit} or \texttt{.module} extension). All following measurements are cumulated per system for each entry point.

\begin{table*}[t]
\centering 
	\begin{tabular}{lrp{4cm}rlrr}
	\toprule
	\textbf{System} & \textbf{Version} & \textbf{Classification} & \textbf{SLOC} & \textbf{\#files} & \textbf{\#OCs} & \textbf{\#includes}
	\\
	\midrule
	AddressBook & 8.2.5.2 & 239 & 51,907 & Adress Administration  & XXX & XXX\\
	SchoolMate & 1.5.4 & 65 & 8,118 & School Administration  & XXX & XXX\\
	TimeClock & 1.04 & 63 & 20,800 & XXX  & XXX & XXX\\
	WebChess & 1.0.0 & 28 & 5,219 & XXX  & XXX & XXX\\
	\midrule
	Drupal & 7.5.0 & 125 & 52,464 & Content Management System  & XXX & XXX\\
	phpBB & 3.1.9 & 1,398 & 327,371 & Bulletin Board  & XXX & XXX\\
	phpMyAdmin & 4.6.3 & 871 & 303,582 & Database Administration  & XXX & XXX\\
	\midrule
	Anchor & 0.12.1 & 201 & 15,054 & Content Management System  & XXX & XXX\\
	Kirby & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
	Automad & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
	Monstra & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
	Nibbleblog & XXX & XXX & XXX & Content Management System  & XXX & XXX\\
	\bottomrule
	\end{tabular}
	\caption{Corpus of twelve PHP systems. The file count includes files with a .php,
	.inc, .bit or .module extension.}
	\label{corpus}
\end{table*}

@@

\subsubsection{What literals did we miss?}\label{WhatLiteralsDidWeMiss?}
Our initial approach to understand missed output candidates is to find out whether their surrounding program code was not accessed/accessible, and if so, why. Of special interest are those output candidates that are located in HTML files (possibly with nested PHP), or located in a function definition. Plain HTML files do not require any further execution (of course, unless PHP scripts are embedded) and represent output candidates that just need to be included properly to be part of the output. Aside, output candidates that are part of a function definition are only missed if the corresponding function is never called at runtime.

We started classifying missed output candidates by their string literal context. As the classification statistics in Figure \ref{coverage} illustrate, for almost all systems (except for \textsf{AddressBook} and \textsf{TimeClock}) not-reached output candidates had a function context. Note that \textsf{SchoolMate} was excluded from the diagram since our analysis reached all output candidates for that system.

@@

Finally, we measured for include expressions the overall reach coverage and the overall resolution success rate per system. As illustrated in Figure \ref{fig:include_coverage_results}, except for the small-scale systems and two recent small-scale systems we did not even really reach a great portion of include expressions, although for most systems the resolution success rate was over 80 percent. Nevertheless, one failed attempt to include a script file can result in a cascade of missing more include expressions as the include target script is not executed.

@@

    \caption{Include coverage and accuracy.\label{fig:include_coverage_results}}
\end{figure}

@@

The callgraph snippets in Figure \ref{fig:callgraph} illustrate the idea of detecting the functions’ access points: The nodes at the top depict callback candidates, the nodes at the bottom with labels X, Y, and Z depict function definitions containing not-reached output candidates. In Figure \ref{fig:callgraph} function X can only be accessed through callback candidates, as for Figure \ref{fig:callgraph} function Y can be partially accessed by callback candidates, whereas function Z is never accessible through callback candidates.

@@

For failed function calls to functions containing not-reached output candidates we measured how many of those functions were only, or partially accessibly through indirect calling mechanisms (see Section \ref{sec:inaccessible}). As Figure \ref{fig:output_candidate_explanation} illustrates, around 80 percent (except for \textsf{TimeClock}) all missed output candidates of each system were only accessible through callback candidates, i.e., functions that are only accessible through indirect calling mechanisms.
